---
title: "Exploratory Data Analysis on Bike Buyers"
author: "Chaitanya"
date: "2024-02-16"
output: html_document
---




Import the data from csv file and read the data

```{r}
setwd("C:\\Users\\urooj\\OneDrive\\Desktop\\Masters\\Spring\\R Sessions")
bikebuyers=read.csv("bike_buyers.csv")
head(bikebuyers)
```

Get high level overview of data

```{r}

str(bikebuyers)
```

Apply Descriptive Statistics on the data

```{r}
summary(bikebuyers)
```
If we observe, the categorical columns like "Marital.Status", "Gender", "Education", "Occupation", "Home.Owner", "Commute.Distance", "Region", "Purchased.Bike" aren't displaying any information as they are in character datatype.

Inorder to make them in a readable format, "as.factor" helps converts variable into a factor.

  + Factorization:

    + Converts the variable into a factor data type, which stores unique categories and assigns integer codes to them.
    + Preserves the original order of categories and their relationships.
    + More memory-efficient for variables with many categories.

```{r}
bikebuyers$Marital.Status <- as.factor(bikebuyers$Marital.Status)
bikebuyers$Gender <- as.factor(bikebuyers$Gender)
bikebuyers$Education <- as.factor(bikebuyers$Education)
bikebuyers$Occupation <- as.factor(bikebuyers$Occupation)
bikebuyers$Home.Owner <- as.factor(bikebuyers$Home.Owner)
bikebuyers$Commute.Distance <- as.factor(bikebuyers$Commute.Distance)
bikebuyers$Region <- as.factor(bikebuyers$Region)
bikebuyers$Purchased.Bike <- as.factor(bikebuyers$Purchased.Bike)
```

Now, by changing all the variables into factors, we can view summary of those columns which can be helpful to explore the data.

```{r}
summary(bikebuyers)
```


```{r}
class(bikebuyers)
```

```{r}
View(bikebuyers)
```

Do you need to adjust anything at this point?data types, structure of the data

*Why we use "dplyr"?*
The dplyr package offers several advantages for data manipulation in R.

```{r}
library("dplyr")
bikebuyers <- select(bikebuyers,-ID)
summary(bikebuyers)
```

### Examine Univariate data

Univariate data means examining the samples of one variable.

  + It doesn't concerned with "why". It just describe the data.
  

### Discrete variables vs Continuous variables

 + *Discrete* - It has set of limited values, an example is level of education in the data set.
 
 + *Continuous* - It can be just any number.
 
 
### Two things to discover in EDA are

  + Central tendency - What is the common value? What's the value around which the data is centered?
  
  + Spread - how varied are the data points?
  
*Types of charts we use*

  + Box plot
  
  + Histogram
  
  + Density Plot
  
  + Pie Charts



*Measure of Central tendency for "Income" - continuous data*

```{r}
summary(bikebuyers$Income)
boxplot(bikebuyers$Income)
```

The above box plot indicates the data is Right Skewed with 3 outliers. The Mean and Median aren't same.

The spread of the data can be seen in the histogram.


```{r}
hist(bikebuyers$Income)
```


We use Density plot to see how frequently I can see the data points.


```{r}
plot(density(bikebuyers$Income),main="income density spread")

#?density

```

Exploring categorical variable "Education"

```{r}

summary(bikebuyers$Education)
plot(bikebuyers$Education)

```


Exploring categorical variable "Marital.Status"

```{r}

summary(bikebuyers$Marital.Status)
plot(bikebuyers$Marital.Status)
pie(table(bikebuyers$Marital.Status), main = "Married vs Single Bike Buyers")

```

Exploring "Children"

```{r}
summary(bikebuyers$Children)
plot(bikebuyers$Children)

```
The above scatter plot is difficult to intrepret.

Ask yourself if this is a right data type to intrepret.

Change to factor
```{r}
bikebuyers$Children=factor(bikebuyers$Children)
summary(bikebuyers$Children)
```

```{r}
plot(bikebuyers$Children,xlab="No. of Children",ylab="frequency")
```

After we change the data type we can read the data clearly.


What about those who bought a bike specifically
```{r}
library("dplyr")
bought=filter(bikebuyers, Purchased.Bike== "Yes")
plot(bought$Children)
```
Mini Challenge:
1. What is the most common commute distance? Visualize it?
2. Did more people buy bikes or not? Visualize it?






```{r}
summary(bikebuyers$Commute.Distance)
plot(bikebuyers$Commute.Distance)
```

```{r}
summary(bikebuyers$Purchased.Bike)
plot(bikebuyers$Purchased.Bike)
```


